Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
Cleaning up introduction. This belongs to later sections, if anywhere.

Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States. Both votes were decided by the slimmest of margins, but more significantly, both votes also had poor turnout. The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community. Hence, there have been major efforts towards increasing these attributes within the community. We believe this is a fallacy. Democracy need not suffer under low engagement — in fact, our goal should be to make democracy work with minimal engagement. The ideal governance system is one where citizens do not have to be constantly involved, but the system is still acting as the citizens would like it to.
We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention? We formulated a novel ‘statistical quorum’, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic. We have developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks. We found that the community was able to use the web application to simultaneously improve the legitimacy and efficiency of their decision-making process.
Our prototyped web application for collective decision-making contrasts with existing similar solutions. Our web application integrates new ideas in how to improve this process: use of \textbf{range voting}, a \textbf{dynamic statistical quorum}, and \textbf{voter delegation}. The broad idea is that, instead of just trying to move the voting process online — something many others have done previously — we in fact wish to add value by voting online. Specifically, we can compute results faster, more accurately, and securely, we can use powerful statistical methods to estimate important quantities for the entire voting population, we can allow for members to entrust their vote to others who have more expertise on specific topics than they themselves do, and we can promote clarity of thought and complete understanding among all members for all voting items.
\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1. Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their positions in a highly informative and precise manner. Moreover, if a new proposal is added to a set of existing proposals on which voting was already open, only the new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of each proposal would need reassessing.
A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions. For example, in a community of 100 members, if 20 members all vote in full support of a proposal, then we can conclude with very high confidence that, had all 100 members actually voted, a majority of them would support the proposal (despite in actuality 80 members having not yet voted). On the other hand, if 80 members vote with 41 voting in favor of the proposal and 39 voting in opposition, then the community cannot yet conclude anything with high confidence (despite the high voter turnout). Hence, the suggested course of action would be for the community to gather more votes and continue discussing the proposal. By utilizing a dynamic statistical quorum, we can direct attention and participation within the community towards contentious proposals, while reducing wasted attention directed towards proposals that already have overwhelming support.
\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate. Not all members of a community can participate at all times and not all members have the time to educate themselves about each and every decision being made. By allowing members to define delegates to whom they can give their own vote to — likely those who share their values — their proxy vote can be inferred from their delegates’ votes. In this way, their values are still voiced even if they do not vote directly. If at any point a member decides to vote themselves, then their direct vote is used instead.
All of the above features are combined into a guided process to help community members participate in decision-making. The process starts with a new discussion item where members can comment and discuss a proposal. Moderators then summarize the discussion into three types of points: “in favor,” “against,” and “other”. Members can then use those summaries to help form an opinion before they vote, without having to read every discussion comment made in full detail, thus lowering the barrier of entry into any discussion. As members vote, the app shows the community the average score and confidence in real-time. When the confidence passes a preset threshold, the proposal can be firmly accepted by the community.

>>>>

%Two of the most shocking democratic votes in recent history occurred in 2016: The referendum vote for Britain to leave the European Union (“Brexit”) and the election of Donald Trump as President of the United States. Both votes were decided by the slimmest of margins, but more significantly, both votes also had poor turnout. The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community. Hence, there have been major efforts towards increasing these attributes within the community. We believe this is a fallacy. Democracy need not suffer under low engagement — in fact, our goal should be to make democracy work with minimal engagement. The ideal governance system is one where citizens do not have to be constantly involved, but the system is still acting as the citizens would like it to.
%We explore a fundamentally different approach to making democracy work: \textit{For a fixed level of community attention, how can we optimize the allocation of this attention to maximize productivity?} Can we increase the credibility of decisions made within a democracy while simultaneously skirting the need for an increase in attention? We formulated a novel ‘statistical quorum’, derived using first principles, as one metric the community can use to focus attention, in turn making the entire decision-making process dynamic. We have developed a web application centered around this metric and tested it in a 140-person community that holds weekly councils over the course of 16 weeks. We found that the community was able to use the web application to simultaneously improve the legitimacy and efficiency of their decision-making process.
%Our prototyped web application for collective decision-making contrasts with existing similar solutions. Our web application integrates new ideas in how to improve this process: use of \textbf{range voting}, a \textbf{dynamic statistical quorum}, and \textbf{voter delegation}. The broad idea is that, instead of just trying to move the voting process online — something many others have done previously — we in fact wish to add value by voting online. Specifically, we can compute results faster, more accurately, and securely, we can use powerful statistical methods to estimate important quantities for the entire voting population, we can allow for members to entrust their vote to others who have more expertise on specific topics than they themselves do, and we can promote clarity of thought and complete understanding among all members for all voting items.
%\textit{Range voting} is voting where members of a community express their positions on open proposals by assigning each a score between –1 and +1. Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their positions in a highly informative and precise manner. Moreover, if a new proposal is added to a set of existing proposals on which voting was already open, only the new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of each proposal would need reassessing.
%A \textit{dynamic statistical quorum} uses statistical hypothesis testing to compute the minimum number of required votes needed to achieve a desired level of confidence before a community adopts an outcome as being truly representative of its members’ positions. For example, in a community of 100 members, if 20 members all vote in full support of a proposal, then we can conclude with very high confidence that, had all 100 members actually voted, a majority of them would support the proposal (despite in actuality 80 members having not yet voted). On the other hand, if 80 members vote with 41 voting in favor of the proposal and 39 voting in opposition, then the community cannot yet conclude anything with high confidence (despite the high voter turnout). Hence, the suggested course of action would be for the community to gather more votes and continue discussing the proposal. By utilizing a dynamic statistical quorum, we can direct attention and participation within the community towards contentious proposals, while reducing wasted attention directed towards proposals that already have overwhelming support.
%\textit{Voter delegation} addresses two questionable assumptions about current democratic systems: (1) participants have the time to participate and (2) participants have the knowledge to participate. Not all members of a community can participate at all times and not all members have the time to educate themselves about each and every decision being made. By allowing members to define delegates to whom they can give their own vote to — likely those who share their values — their proxy vote can be inferred from their delegates’ votes. In this way, their values are still voiced even if they do not vote directly. If at any point a member decides to vote themselves, then their direct vote is used instead.
%All of the above features are combined into a guided process to help community members participate in decision-making. The process starts with a new discussion item where members can comment and discuss a proposal. Moderators then summarize the discussion into three types of points: “in favor,” “against,” and “other”. Members can then use those summaries to help form an opinion before they vote, without having to read every discussion comment made in full detail, thus lowering the barrier of entry into any discussion. As members vote, the app shows the community the average score and confidence in real-time. When the confidence passes a preset threshold, the proposal can be firmly accepted by the community.
